---
title: "Day 3"
author: "Raghu Chandramohan"
date: "6/13/2018"
output: 
  html_document: 
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidyverse
```{r include=FALSE}
library(tidyverse)
library(magrittr)
```

## General Exercise

### Pipes
Rewrite the following code chunks as pipes (magrittr):

#### Chunk 1
```
my_cars <- mtcars[, c(1:4, 10)]
my_cars <- my_cars[my_cars$disp > mean(my_cars$disp), ]
my_cars <- colMeans(my_cars)
```
```{r}
#My solution
my_cars = as.tibble(mtcars[,c(1:4,10)]) %>% filter(disp > mean(disp)) %>% map(mean)
# Answer
mtcars %>% 
  select(c(1:4, 10)) %>% 
  filter(disp > mean(disp)) %>% 
  colMeans() -> my_cars
```

#### Chunk 2
The `summary(x)` function is a bit special: when you type `summary(x)` in the console, `print` is called in an implicite way. Pipe call does not do, so you will have to invoke `print` in an explicite way. But the `%T>%` does unbranch for one call only, you will have to make printing of the `summary` a one single composed call using `{}`.

```
summary(cars)
colSums(cars)
```
```{r}
#My solution
cars %T>% {print(summary(.))} %>% colSums()
```

#### Chunk 3
```
cor(mtcars$gear, mtcars$mpg)
cor(mtcars)
```
```{r}
mtcars %$% cor(gear, mpg)
mtcars %>% cor()
#or
mtcars %>% cor(.,.)
```

#### Chunk 4
Given is the `dim_summary(nrows, ncols)` function which takes matrix `nrows` and `ncols` as arguments and prints this info about its dimensions:
```{r}
dim_summary <- function(nrows, ncols) {
  print(
    paste0('Matrix M has: ', nrows, ' rows and ', ncols, ' columns.')
  )
}
```
Rewrite the code chunks below as pipes:
```
distr1 <- rnorm(16)
M <- matrix(distr1, ncol = 4)
plot(M)
M <- M + sample(M)
dim_summary(nrows = nrow(M), ncols = ncol(M))

```
```
distr2 <- rnorm(16)
N <- matrix(distr2, ncol = 4)
colnames(N) = letters[1:4]
summary(N)
```

```
P <- M %x% t(N)
heatmap(P)
colnames(P) <- letters[1:dim(P)[2]]
cor(P[ ,'a'], P[ ,'i'])
```

```{r}
M = rnorm(16) %>% matrix(ncol = 4) %T>% plot() %>% `+`(.,sample(.)) %T>% {dim_summary(nrows = nrow(.), ncols = ncol(.))}

N = rnorm(16)  %>% matrix(ncol = 4) %>% `colnames<-`(letters[1:4]) %T>% {print(summary(.))}

M %x% t(N) %T>% 
  heatmap() %>% `colnames<-`(letters[1:dim(.)[2]]) %>%
  as.tibble() %$%
  cor(a,i)
```

## Tibbles

### Task 1
- Convert the `mtcars` dataset to a tibble `vehicles`.
  -  Select the number of cylinders (`cyl`) variable using:
  -  the `[[index]]` accessor,
  -  the `[[string]]` accessor,
  -  the `$` accessor.
- Do the same selection as above, but using pipe and placeholders (use all thre ways of accessing a variable).
- Print the tibble.
- Print the 30 first rows of the tibble.
- Change the default behaviour of printing a tibble so that at least 15 and at most 30 rows are printed.
- What is the difference between the `tibble.print_max` and `dplyr.print_min`? Is there any? Test it.
- Convert `vehicles` back to a `data.frame` called `automobiles`.
```{r}
vehicles = as.tibble(mtcars)
vehicles[[2]]
vehicles[["cyl"]]
vehicles$cyl
```
```{r}
X = vehicles %>% 
  select(cyl) 
X %>% head(30)
options(tibble.print_min = 15, tibble.print_max = 30)

automobiles = vehicles %>% as.data.frame()
```

### Task 2
```{r}
tribble(
  ~id, ~event, ~data,
  1,	"success",	"24-04-2017",
  2,	"failed",	"25-04-2017",
  3,	"failed",	"25-04-2017",
  4,	"success",	"27-04-2017")
```

### Task 3
Compare the performance of `as.data.frame()`, `as_data_frame()` and `as_tibble()` on a 100 x 30 matrix filled with random integers. Use package `microbenchmark`. Fill in your result here in the Tidyverse Lab sheet, Tibbles – performance.
```{r}
library(microbenchmark)
X = sample(1:10, 100*30, replace = T) %>% matrix(ncol = 30)
microbenchmark(as.data.frame(X))
microbenchmark(as_data_frame(X))
microbenchmark(as.tibble(X))
microbenchmark(as_tibble(X))
```

### Task 4
Do you think tibbles are lazy? Try to create a tibble that tests whether lazy evaluation applies to tibbles too.
```{r}
tibble(x = sample(1:10, size = 10, replace = T), y = log10(x))
```

## Parsing
Parse the following vectors:

- vec1 <- c(1, 7.2, 3.84, -5.23) – parse it as double.
- now parse the same vector c(1, 7.2, 3.84, -5.23) as integer. What happens?
- Can you still parse it as integer somehow?
- Parse as double vec2 <- c('2', '3,45', '?', '-7,28')
- Parse correctly vec3 <- c('2', '3,45', '?', '-7.28')
- Parse the following guessing the parser:  vec4 <- c('barrel: 432.7$', 'liter: 15.42PLN', 'gallon costs approx 32.1SEK', 'sunny, wind gusts up till 55m/s')
- Can you parse vec4 as number? Do it if you can.
- Parse vec5 <- "25 Dec 2015" as date (hint: ?parse_date()).
- Parse 10_Jul_1410 as date.
```{r}
vec1 = c(1, 7.2, 3.84, -5.23)
vec2 = c('2', '3,45', '?', '-7,28')
vec3 = c('2', '3,45', '?', '-7.28')
vec4 = c('barrel: 432.7$', 'liter: 15.42PLN', 'gallon costs approx 32.1SEK', 'sunny, wind gusts up till 55m/s')
vec5 = "25 Dec 2015"

parse_double(vec1)
parse_integer(vec1)
parse_integer(as.integer(vec1))
parse_double(vec2, na = "?", locale = locale(decimal_mark = ","))
parse_number(vec2, na = "?", locale = locale(decimal_mark = ","))
guess_parser(vec4)
parse_guess(vec4)
parse_date(vec5, format = "%d %b %Y")
parse_date("10_Jul_1410", format = "%d_%b_%Y")
```

## NYC flights challenge
The `nycflights13` package contains information about all flights that departed from NYC (i.e., EWR, JFK and LGA) in 2013: 336,776 flights with 16 variables. To help understand what causes delays, it also includes a number of other useful datasets: weather, planes, airports, airlines. We will use it to train working with tibbles and `dplyr`.

```{r include=FALSE}
library(nycflights13)
library(dplyr)
```

### Task 1: Selecting column
load the nycflights13 package (install if necessary),
read about the data in the package docs,
inspect the flights tibble.
select all columns but carrier and arr_time,
select carrier, tailnum and origin,
hide columns from day through carrier,
select all columns that have to do with arrival (hint: ?tidyselect),
select columns based on a vector v <- c("arr_time", "sched_arr_time", "arr_delay"),
rename column dest to destination using:
select() and
rename()
What is the difference between the two approaches?
```{r}
flights %>% select(-c(carrier, arr_time)) 
flights %>% select(carrier, tailnum, origin) 
flights %>% select(-(day:carrier)) 
flights %>% select(contains("arr_")) 

v <- c("arr_time", "sched_arr_time", "arr_delay")
flights %>% select(v)

flights %>% select(destination = dest, everything())
flights %>% rename(destination = dest)
```

### Task 2: Filtering rows
filter only the flights that arrived ahead of schedule,
filter the flights that had departure delay between 10 and 33,
fish out all flights with unknown arrival time,
retrieve rows 1234:1258 (hint: ?slice),
sample (?sample_n()) 3 random flights per day in March,
show 5 most departure-delayed flights in January per carrier,
retrieve all unique() routes and sort them by origin,
retrieve all distinct() routes and sort them by origin,
is unique() more efficient than distinct()?

```{r}
flights %>% filter(arr_delay < 0)

flights %>% filter(dep_delay >= 10 & dep_delay <= 30)
flights %>% filter(between(dep_delay,10,30)) #alternative

flights %>% filter(is.na(arr_time))

flights %>% slice(1234:1258)

flights %>% filter(near(3, month)) %>% group_by(day) %>% sample_n(3)

flights %>% filter(month == 1) %>% group_by(carrier) %>% top_n(5, dep_delay)

flights %>% mutate(routes = paste0(origin , "_" , dest)) %>% distinct(routes) %>% arrange(routes)

flights %>% mutate(routes = paste0(origin , "_" , dest)) %>% {unique((.)["routes"])} %>% arrange(routes)
```

### Task 3: Trans(mutations)
air_time is the amount of time in minutes spent in the air. Add a new column air_spd that will contain aircraft’s airspeed in mph,

as above, but keep only the new air_spd variable,

use rownames_to_column() on mtcars to add car model as an extra column,

```{r}
flights %>% mutate(air_spd = distance/(air_time/60))
flights %>% transmute(air_spd = distance/(air_time/60))

mtcars %>% rownames_to_column()
```


### Task 4: Groups and Counts
use group_by(), summarise() and n() to see how many planes were delayed (departure) every month,
```{r}
flights %>% filter(dep_delay > 0) %>% group_by(month) %>% summarise(count_dep_delay = n())
```

do the same but using tally() and count(),
```{r}
flights %>% filter(dep_delay > 0) %>% group_by(month) %>% tally()
flights %>% filter(dep_delay > 0) %>% count(month)
```
what was the mean dep_delay per month?
```{r}
flights %>% group_by(month) %>% summarise(mean(dep_delay, na.rm = T))
```

count the number of incoming delayed flights from each unique origin and sort origins by this count (descending),
```{r}
flights %>% filter(arr_delay > 0) %>% group_by(origin) %>% summarise(count = n()) %>% arrange(desc(count))
```

do the same using tally()
```{r}
flights %>% filter(arr_delay > 0) %>% group_by(origin) %>% tally(sort = T)
```

use summarise() to sum total dep_delay per month in hours,
```{r}
flights %>% group_by(month) %>% summarise(sum_dep_delay = sum(dep_delay/60, na.rm = T))
```

use the wt parameter of count() (works with tally() too) to achieve the same,
```{r}
flights %>% group_by(month) %>% count(wt = dep_delay/60)
```

run group_size() on carrier what does it return?
```{r}
flights %>% group_by(carrier) %>% group_size()
```

use n_groups() to check the number of unique origin-carrier pairs,
```{r}
flights %>% group_by(carrier) %>% n_groups()
```

```
flights %>% 
  group_by(origin) %>% 
  mutate(mean_delay_orig = (mean(dep_delay, na.rm = T) + mean(arr_delay, na.rm = T)) / 2) %>%
  ungroup() %>%
  group_by(carrier) %>%
  mutate(mean_delay_carr = (mean(dep_delay, na.rm = T) + mean(arr_delay, na.rm = T)) / 2) %>%
  select(origin, carrier, mean_delay_orig, mean_delay_carr)
```

### Task 5: Joins
```{r}
set1 = tribble(~id,	~color,
"id1",	"grey",
"id1",	"red",
"id2",	"green",
"id3",	"blue")

set2 = tribble(~id,	~color,
"id1",	"grey",
"id1",	"red",
"id2",	"green",
"id3",	"blue")

left_join(set1, set2)
right_join(set1, set2)
inner_join(set1, set2)
full_join(set1, set2)
anti_join(set1, set2)

```

## Tidying data
Now time to do some data tidying. First install a package with some untidy data:
```{r}
library(EDAWR)
EDAWR::cases %>% gather(`2011`, `2012`, `2013`, key = "year", value = "value")
```

now time for the pollution dataset. Tidy it so that there separate columns for large and small pollution values.
```{r}
pollution %>% spread(key = size, value = amount)
```

the storms dataset contains the date column. Make it into 3 columns: year, month and day. Store the result as tidy_storms
```{r}
x = storms %>% separate(date, into = c("Year", "Month", "Day"), sep = "-", convert = T)
```

now, merge year, month and day in tidy_storms into a date column again but in the “DD/MM/YYYY” format.
```{r}
x %>% unite(col = Date, Year, Month, Day, sep = "/" )
```

# ggplot


# ggmap
As the name suggests, the get_map function from ggmap is used to get maps. The maps are of the class ggmap and you can get many different maptypes.

Use ggmap(), the function, to plot a ggmap object created by get_map and try out a few maptypes.
Note: Unless you have adjusted ggmap settings with a google API key the get_map seems to be a bit unreliable during testing. If the command fails, try again a few times. Save it to a variable so you don’t have to keep querying the map servers.

```{r}
library(ggmap)
map_obj = get_map(location = "Houston", zoom = 10)
ggmap(map_obj)
```

Trying out some other maptypes:
```{r}
map_obj_water = get_map(location = "Houston", zoom = 10, maptype = "watercolor")
ggmap(map_obj_water)

map_obj_sat = get_map(location = "Houston", zoom = 10, maptype = "satellite")
ggmap(map_obj_sat)

map_obj_ton = get_map(location = "Houston", zoom = 10, maptype = "toner-lines")
map_ton = ggmap(map_obj_ton)
```

## Add data to the map
Being able to output maps is great but, as with everything in life, it becomes more interesting when you can overlay it with data.

### Points | geocode() and geom_point()
ggmap::geocode() is a nifty function that returns the latitude and longitude of a location. Use this to get the geocode of a place within the map that you created earlier in the exercise. Your favourite gym or restaurant for example?

In my example code, I show how to use tibble to get geocodes for several locations and bind them together to a data.frame.

```{r}
raghu_locations = tibble(location = c("Baylor College of Medicine, Houston"))
raghu_geo = geocode(raghu_locations$location)
raghu_places = cbind(raghu_locations, raghu_geo)
```

When you have your locations dataset ready, use geom_point to add markers to your map for your locations. Feel free to use geom_text to also add labels for your locations. Notice that you can treat the ggmap function like ggplot.
```{r}
ggmap(map_obj_water) + 
  geom_point(data = raghu_places, aes(x = lon, y = lat), color = "black", size = 3) +
  geom_text(data = raghu_places, aes(label = location), hjust = 0, vjust = -1)
```

### Routes | trek() and geom_path()
trek is a function from ggmap that takes a start location and a stop location and gives a list of latitude and longitude coordinates connecting the two. You can give several different modes of transit, but why anyone would choose anything other than bicycling is beyond me.

Combining trek and ggplot2 geom_path we can plot a route between two places, as long as one exists. Regrettably it doesn’t seem like airplane transits are included yet. Read the help page for trek and try it out.

Note: Make sure your background map covers both start and stop position!
```{r}
houston_zoomout = get_map(c(lon = -96.5, lat = 29.8), zoom = 8, maptype = "watercolor", scale = 2)
hou_aus = trek("Houston", "Austin", structure = "route", mode = "driving")
ggmap(houston_zoomout) +
  geom_path(aes(x= lon, y = lat), color = "blue", data = hou_aus, size = 2, alpha = 0.5, lineend = "round")
```

### Pubs
What do they mean by pubs? Do they mean publications? Nope! We are going to plot actual pubs, bars, inns in the UK using a public dataset from https://www.getthedata.com/open-pubs.

Download the dataset pubs.rda from course materials.

Note: I have processed this data slightly, to see how unhide the following code box.

```{r}
load("/Users/raghu/Downloads/pubs.rda")
pubs_tb = as_tibble(pubs)
pubs_cambridge = pubs_tb %>% filter(city == "Cambridge")

cambrige_map = get_map("Cambrige, UK", zoom = 15, maptype = "toner-lines")
ggmap(cambrige_map) +
  geom_point(data = pubs_cambridge, aes(x = longitude, y = latitude), color = "red")

ggmap(cambrige_map) +
  geom_density_2d(data = pubs_cambridge, aes(x = longitude, y = latitude)) +     stat_density_2d(data = pubs_cambridge, aes(x = longitude, y = latitude,
    fill = ..level..,alpha = ..level..), geom = "polygon", color = NA)
```

